---
title: "EDA"
author: "Allison Randy-Cofie"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data
```{r}
library(tidyverse)
library(patchwork)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```

## `group_by`
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup()
```


## counting things
count month observations
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n()) #name the variable n_obs
```

we can use `count()`
```{r}
weather_df %>% 
  count(month, name = "n_obs")
```

**NEVER** use base R's `table` function
```{r, eval=FALSE}
weather_df %>% 
  pull(month) %>% 
  table() #this result is not a data frame
```


Other helpful counters
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n(),
            n_days = n_distinct(date))
```


## A digression on 2x2 tables
```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~"cold",
      tmax >=5 ~ "not_cold",
      TRUE ~" "
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count=n())
```


```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~"cold",
      tmax >=5 ~ "not_cold",
      TRUE ~" "
    )
  ) %>% 
  janitor::tabyl(name,cold)
```


## General summaries

You can do lots of summaries

```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarize(
    mean_tmax=mean(tmax, na.rm=TRUE),
    mean_prcp=mean(prcp,na.rm=TRUE),
    median_tmin=median(tmin,na.rm=TRUE)
  )
```


This is a dataframe! So we can make a plot
```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarize(
    mean_tmax=mean(tmax, na.rm=TRUE),
    mean_prcp=mean(prcp,na.rm=TRUE),
    median_tmin=median(tmin,na.rm=TRUE)
  ) %>% 
  ggplot(aes(x=month,y=mean_tmax, color=name))+
  geom_point()+
  geom_line()
```


Suppose you want to summarize many columns.
```{r}
weather_df %>% 
  group_by(name,month) %>% 
  summarize(across(prcp:tmin,mean))
```

Reminder: sometimes your results are easier to read in another format.

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from=name,
        values_from=mean_tmax) %>% 
  knitr::kable(digits = 1)
  
```

## `group_by` and `mutate`

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm=TRUE),
    centered_tmax = tmax-mean_tmax
  ) %>% 
  ggplot(aes(x=date,y=centered_tmax,color=name)) +
  geom_point()
```


what about window functions
```{r}
weather_df %>% 
  group_by(name,month) %>% 
  mutate(
    temp_rank=min_rank(tmax)
  ) %>% 
  filter(temp_rank==1)
```

## lag
```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = lag(tmax)) #opposite is lead
```


## Quick note
Summarize only gets you so far.

