---
title: "viz_part_01"
author: "Allison Randy-Cofie"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's import data
```{r}
library(tidyverse)
library(ggridges)
```

This is code from the course website

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Let's make a scatterplot

```{r}
ggplot(weather_df, aes(x=tmin, y=tmax)) +
  geom_point()
```

Let's make the same scatterplot, but different

```{r}
weather_df %>% 
  drop_na() %>% 
  #filter(name == "CentralPark_NY") %>%  this shows only NY
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point()
```

Let's keep making the same plot but different.

```{r}
weather_scatterplot = 
  weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x=tmin,y=tmax))

weather_scatterplot +
  geom_point()
```

## Let's fancy this up a bit
```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point() 
 # geom_smooth() applies a line for each location instead of the bottom one
```

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point(alpha=0.3) + #makes the points transparent
 geom_smooth(se = FALSE)  #gets rid of error bars
```



```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax)) +
  geom_point(aes(color=name)) +
  geom_smooth() #runs a line through the middle of the dataset
```

Maybe make separate panels.

```{r}
weather_df %>% 
  ggplot(aes(x=tmin,y=tmax,color=name)) +
  geom_point(alpha=0.3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) #makes different panels without excessive code name~ . will put names on the rows. so it is row ~ column
```

`tmax` vs `tmin` is boring, let's spice it up some.

```{r}
weather_df %>% 
  ggplot(aes(x=date, y=tmax, size = prcp)) +
  geom_point(alpha=0.3) +
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name)
```



```{r}
weather_df %>% 
  ggplot(aes(x=date, y=tmax, color = name)) +
  geom_point(aes(size = prcp), alpha=0.3) +
  geom_smooth(se = FALSE)+
  facet_grid(. ~ name) +
  theme(axis.text.x=element_text(angle = 90, vjust = 9.5, hjust=1)) # rotates the x axis labels
```


## Some quick stuff
```{r}
weather_df %>% 
 ggplot(aes(x=tmin, y=tmax)) +
  geom_hex()
```

Histograms, densityplots, boxplots and violins,,.

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, color=name)) +
  geom_histogram()

```

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name)) +
  geom_histogram()
```

```{r}
weather_df %>% 
  ggplot(aes(x=tmin, fill=name)) +
  geom_histogram()+
  facet_grid(. ~ name)
```


Let's see more options!
```{r}
weather_df %>% 
  ggplot(aes(x=tmax, color=name)) +
  geom_density()
```

```{r}
weather_df %>% 
  ggplot(aes(x=tmax, fill=name)) +
  geom_density()
```

```{r}
weather_df %>% 
  ggplot(aes(x=tmax, fill=name)) +
  geom_density(alpha=.3)
```
 
 Boxplots?
```{r}
weather_df %>% 
  ggplot(aes(y = tmax)) +
  geom_boxplot()
```
 
```{r}
weather_df %>% 
  ggplot(aes(x=name, y = tmax)) +
  geom_boxplot()
```
 
 
```{r}
weather_df %>% 
  ggplot(aes(x=name, y = tmax, fill=name)) +
  geom_boxplot()
```
 
 
 Violin plots come in handy when you have like 20/30 things you are paneling by
```{r}
weather_df %>% 
  ggplot(aes(x=name, y = tmax, fill=name)) +
  geom_violin()
```
 
OR

```{r}
weather_df %>% 
  ggplot(aes(x=tmax,y=name)) +
  geom_density_ridges()
```

## Saving and embedding plots
First -- let's save a plot.
```{r}
weather_scatterplot=
  weather_df %>% 
  ggplot(aes(x=date, y=tmax, color=name)) +
  geom_point(aes(size=prcp),alpha=0.3) +
  geom_smooth(se=FALSE) +
  facet_grid(. ~ name)

weather_scatterplot

ggsave(
  file="results/weather_scatterplot.pdf", 
  plot=weather_scatterplot, 
  width = 8, height = 5)
```


```{r, fig.width=10}
weather_scatterplot
```

